\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
Every book needs an author. My life contains pages of many types; some are blurry, some are crumpled, some are torn, some are black and white, some are colorful, some are stunning, some are full, and some are empty. I can only be thankful to my Author when I reread them. Though the pages looked like a randomly assembled collage at some point, they make sense together today. Through His mercy and love, I trust that the empty pages will be filled with everlasting words and art. 

\medskip

Every pupil needs a master, a role model, someone who started the journey first and is willing to guide, share knowledge, encourage, and correct. I was fortunate enough to have two of them during this adventure. I thank Prof. Dr. Raphael Sznitman for granting me the opportunity to work in his lab under his supervision. For looking at me in full size and not as a dot from the humongous mountain he has climbed. For being humble and kind, for always being willing to help and to discuss, and for patience. I thank Dr. Pablo Márquez-Neila for not letting me swallow ideas without chewing on them first and for always being alert at details that often pointed out my errors. I also thank him for his humility and care for others' processes and deadlines. I thank both for teaching me valuable lessons about research, communication, and life.

\medskip

Every explorer needs a compass to navigate unknown paths. This was especially true during this research journey. I thank my co-advisor, Dr. Damien Teney, for his interest in my process and his willingness to share his perspectives in his field of expertise. His ideas and work inspired me to pursue more profound knowledge and realize the limits of conventional ideas. I thank my mentor, PD Dr. Sigve Haug, for assuming his role with the best disposition. I also owe my thanks to the University of Bern and the Swiss National Science Foundation (SNSF) for their relevant support in funding my research.

\medskip

Every combatant needs a squad. At the AIMI lab, I found a team of brilliant and competent people who selflessly offered me their friendship and support. First, I thank those who, at the beginning, made things easier for me and offered to help, including Vasily Tolkachev, Dr. Mathias Gallardo, Dr. Tatiana Fountoukidou, and Dr. Thomas Kurmann. I thank the other Ph.D. students, Javier Gamazo-Tejero, Lars Doorenbos, Lukas Zbinden, Fei Wu, Paulo Sampaio, and the later-arrived Moritz Schmid. The moments we shared made the process fun and interesting, and I certainly learned something from each of them. My gratitude also goes to all the other members of the team, including Prisca Dotti, Dr. Negin Ghamsarian, Marta Colmenar Herrera, Dr. Alain Jungo, Dr. Theodoros Pissas, Dr. Christopher Hahne, Dr. Corin Otesteanu, Michel Hayoz, Davide Scandella, Tamara Danilovska, and everyone else I had the pleasure of sharing with, even if for a short time. Finally, I thank the administrative staff, Pia Eichholzer, Diana Bethge, and Daniela Schmidt, for always being willing to help.

\medskip

I was lucky to have an extended squad of people who supported me and shared with me in person or from thousands of kilometers away.  I thank Jose Arturo García, William Giraldo, Sergio Castro, Ivan Said Góngora, Eduardo Osorio, Miguel García, Nicolás Camargo, Cassandra Chetty and her lovely family, Andreina Ravani, Denis Borel, Emre Hacan and his wife Gülcan, Yasemin Aversa, Didem Ekrem, Gresa Asani, and my MAIA friends, specially Roa'a Khaled, Valeria Abramova, Tewodros Weldebirhan Arega and Ahmed Gouda. I also extend my thanks to Umut and Gökcen Parlar and their parents, Canan and Turgay, for their kindness, support, and positive attitude.

\medskip

Every tree needs roots. I call my roots family, the people who nurture me and are the foundation of the person I am today. I want to thank my family, including my parents, Cruz Stella and Gustavo, my siblings Natalia and Juan Felipe, and my little niece, Violeta. Their love, prayers, sacrifices, and constant support have been a blessing that distance cannot extinguish. I also thank my aunt Sandra and her husband Pepe, my aunts Libia, Liliana, Neyla, and Cecilia, my uncle Humberto, my cousins Sandra Teresa, Monica, Edward, Iván, and Daniela, and my grandma Teresa.

\medskip

Every heart needs love, someone who is always there, no matter what, who confronts your deepest shadows and is not terrified, who sees in you what other people cannot see. I met Ebru Parlar more than a thousand days ago, and since the beginning, her love, company, support, understanding, care, and cuteness have been priceless treasures in my life. I thank her for believing in me, helping me grow, and being a shiny little star that illuminates my days. 



\end{acknowledgements}